  As \citet[P~1 complete]{Joyce:FW} most eloquently says:
  \begin{quotation}\small
    riverrun, past Eve and Adam's, from swerve of shore to bend of
    bay, brings us by a commodius vicus of recirculation back to Howth
    Castle and Environs.
    
    Sir Tristram, violer d'amores, fr'over the short sea, had
    passencore rearrived from North Armorica on this side the scraggy
    isthmus of Europe Minor to wielderfight his penisolate war: nor
    had topsawyer's rocks by the stream Oconee exaggerated themselse
    to Laurens County's gorgios while they went doublin their mumper
    all the time: nor avoice from afire bellowsed mishe mishe to
    tauftauf thuartpeatrick: not yet, though venissoon after, had a
    kidscad buttended a bland old isaac: not yet, though all's fair in
    vanessy, were sosie sesthers wroth with twone nathandjoe. Rot a
    peck of pa's malt had Jhem or Shen brewed by arclight and rory end
    to the regginbrow was to be seen ringsome on the aquaface.
        
    The fall
    (baba\-badal\-gharagh\-takammi\-narronn\-konn\-bronn\-% break long word
    ton\-ne\-rronn\-tuonn\-thunn\-trovar\-rhoun\-awn\-% break long word
    skawn\-too\-hoo\-hoor\-den\-ent\-hur\-nuk!)\ of
    a once wallstrait oldparr is retaled early in bed and later on
    life down through all christian minstrelsy. The great fall of the
    offwall entailed at such short notice the pftjschute of Finnegan,
    erse solid man, that the humptyhillhead of humself prumptly sends
    an unquiring one well to the west in quest of his tumptytumtoes:
    and their upturnpikepointandplace is at the knock out in the park
    where oranges have been laid to rust upon the green since
    devlinsfirst loved livvy.
  \end{quotation}
  and then later, in the last lines of the book, \citep[P~627,
  Lines 8--16]{Joyce:FW}:
  \begin{quote}\small
    Yes. Carry me along, taddy, like you done through the toy fair! If
    I seen him bearing down on me now under whitespread wings like
    he'd come from Arkangels, I sink I'd die down over his feet,
    humbly dumbly, only to washup. Yes, tid. There's where. First. We
    pass through grass behush the bush to. Whish! A gull. Gulls. Far
    calls. Coming, far! End here.  Us then. Finn, again! Take.
    Bussoftlhee, mememormee! Till thousendsthee. Lps. The keys to.
    Given! A way a lone a last a loved a long the
  \end{quote}
